---
title: "Computational Modeling - Week 4 - Assignment 2 - Part 1"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?
- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results
- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```{r}
library(ggplot2)

####GRID APPROXIMATION 
#Define the grid
dens = 20 
p_grid = seq(from=0, to=1, length.out=dens)

#Define the prior (here --> flat prior, all have prob=1)
prior = rep(1,dens)

#Compute the likelihood of each value in the grid - 3 correct answers out of 6 questions
likelihood = dbinom(3, size = 6, prob=p_grid)

#Compute the posterior - likelihood by prior 
unstd.posterior = likelihood*prior
posterior <- unstd.posterior / sum(unstd.posterior)

#Plot the posterior 
data=data.frame(grid=p_grid,posterior=posterior,prior=prior,likelihood=likelihood)

ggplot(data,aes(grid,posterior))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/dens),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")


#### QUADRATIC APPROXIMATION
install.packages("rethinking")
library(rethinking)

globe.qa <- map( alist( 
  w ~ dbinom(9,p) , # binomial likelihood  
  p ~ dunif(0,1) # uniform prior  
  ), data=list(w=6)) 

#Display summary of quadratic approximation 
precis( globe.qa )


#What is the probability he knows more than chance (0.5)
sum((data$grid > 0.5)*data$posterior)

```

2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.

```{r}
####GRID APPROXIMATION - for each teacher 

df <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

posterior_list = NULL

for (i in 1:4){
  dens = 1000
  p_grid = seq(from=0, to=1, length.out = dens)
  prior = rep(1, dens)
  likelihood = dbinom(df$Correct[i], df$Questions[i], prob=p_grid)
  unstd.posterior = likelihood*prior
  posterior = unstd.posterior / sum(unstd.posterior)
  posterior_list = c(posterior_list, posterior)
}
posterior_list

results = data.frame(RF = posterior_list[1:1000], KT=posterior_list[1001:2000], JS=posterior_list[2001:3000], MW=posterior_list[3001:4000])

#Plot the posterior for each teacher 
results$p_grid = seq(from=0, to=1, length.out = 1000)
results$prior = rep(1, 1000)

RF = ggplot(results,aes(p_grid,RF))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/dens),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

KT = ggplot(results,aes(p_grid,KT))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/dens),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

JS = ggplot(results,aes(p_grid,JS))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/dens),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

MW = ggplot(results,aes(p_grid,MW))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/dens),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

attach(mtcars)
par(mfrow=c(2,2))
RF
KT
JS
MW

```


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r}
df <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

posterior_list_2 = NULL

for (i in 1:4){
  p_grid = seq(from=0, to=1, length.out = 1000)
  prior = dnorm(p_grid, 0.8,0.2)
  likelihood = dbinom(df$Correct[i], df$Questions[i], prob=p_grid)
  unstd.posterior = likelihood*prior
  posterior = unstd.posterior / sum(unstd.posterior)
  posterior_list_2 = c(posterior_list_2, posterior)
}
posterior_list_2

results_2 = data.frame(RF = posterior_list_2[1:1000], KT=posterior_list_2[1001:2000], JS=posterior_list_2[2001:3000], MW=posterior_list_2[3001:4000])

#Plot the posterior for each teacher 
results_2$p_grid = seq(from=0, to=1, length.out = 1000)
results_2$prior = dnorm(seq(from=0, to=1, length.out = 1000), 0.8,0.2)

RF_2 = ggplot(results_2,aes(p_grid,RF))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

KT_2 = ggplot(results_2,aes(p_grid,KT))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

JS_2 = ggplot(results_2,aes(p_grid,JS))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

MW_2 = ggplot(results_2,aes(p_grid,MW))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

attach(mtcars)
par(mfrow=c(2,2))
RF_2
KT_2
JS_2
MW_2
```


4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

```{r}
df2 <- data.frame(
  Correct=c(300,200,16000,6600),
  Questions=c(600,200,19800,13200),
  Teacher=c("RF","KT","JS","MW"))

#UNIFORM PRIOR
posterior_list_3 = NULL

for (i in 1:4){
  dens = 1000
  p_grid = seq(from=0, to=1, length.out = dens)
  prior = rep(1, dens)
  likelihood = dbinom(df2$Correct[i], df2$Questions[i], prob=p_grid)
  unstd.posterior = likelihood*prior
  posterior = unstd.posterior / sum(unstd.posterior)
  posterior_list_3 = c(posterior_list_3, posterior)
}
posterior_list_3

results_3 = data.frame(RF = posterior_list_3[1:1000], KT=posterior_list_3[1001:2000], JS=posterior_list_3[2001:3000], MW=posterior_list_3[3001:4000])

#Plot the posterior for each teacher 
results_3$p_grid = seq(from=0, to=1, length.out = 1000)
results_3$prior = rep(1, 1000)

RF_3 = ggplot(results_3,aes(p_grid,RF))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

KT_3 = ggplot(results_3,aes(p_grid,KT))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

JS_3 = ggplot(results_3,aes(p_grid,JS))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

MW_3 = ggplot(results_3,aes(p_grid,MW))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

attach(mtcars)
par(mfrow=c(2,2))
RF_3
KT_3
JS_3
MW_3

#An attempt to plot them in once 
library(tidyverse)
long_3 = gather(results_3, name, value, RF:MW)

ggplot(long_3,aes(p_grid,value), colour=name)+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")



#----------------------------------------------------------------------------------------


#NORMAL PRIOR
posterior_list_4 = NULL

for (i in 1:4){
  dens = 1000
  p_grid = seq(from=0, to=1, length.out = 1000)
  prior = dnorm(p_grid, 0.8,0.2)
  likelihood = dbinom(df2$Correct[i], df2$Questions[i], prob=p_grid)
  unstd.posterior = likelihood*prior
  posterior = unstd.posterior / sum(unstd.posterior)
  posterior_list_4 = c(posterior_list_4, posterior)
}
posterior_list_4

results_4 = data.frame(RF = posterior_list_4[1:1000], KT=posterior_list_4[1001:2000], JS=posterior_list_4[2001:3000], MW=posterior_list_4[3001:4000])

#Plot the posterior for each teacher 
results_4$p_grid = seq(from=0, to=1, length.out = 1000)
results_4$prior = dnorm(seq(from=0, to=1, length.out = 1000), 0.8,0.2)

RF_4 = ggplot(results_4,aes(p_grid,RF))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

KT_4 = ggplot(results_4,aes(p_grid,KT))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

JS_4 = ggplot(results_4,aes(p_grid,JS))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

MW_4 = ggplot(results_4,aes(p_grid,MW))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

attach(mtcars)
par(mfrow=c(2,2))
RF_4
KT_4
JS_4
MW_4

```


5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

```{r}
#I would use a prior of low probability - maybe 0.5 and with a relatively small standard deviation of 0.1. (to keep it more conservative)
#Because we expect him to be at chance level (it is difficult to always answer wrong)

posterior_list_5 = NULL

for (i in 1:4){
  dens = 1000
  p_grid = seq(from=0, to=1, length.out = dens)
  prior = dnorm(p_grid, 0.5,0.1)
  likelihood = dbinom(df2$Correct[i], df2$Questions[i], prob=p_grid)
  unstd.posterior = likelihood*prior
  posterior = unstd.posterior / sum(unstd.posterior)
  posterior_list_5 = c(posterior_list_5, posterior)
}
posterior_list_5

results_5 = data.frame(RF = posterior_list_5[1:1000], KT=posterior_list_5[1001:2000], JS=posterior_list_5[2001:3000], MW=posterior_list_5[3001:4000])

#Plot the posterior for each teacher 
results_5$p_grid = seq(from=0, to=1, length.out = 1000)
results_5$prior = dnorm(seq(from=0, to=1, length.out = 1000), 0.5,0.1)

RF_5 = ggplot(results_5,aes(p_grid,RF))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

KT_5 = ggplot(results_5,aes(p_grid,KT))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

JS_5 = ggplot(results_5,aes(p_grid,JS))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

MW_5 = ggplot(results_5,aes(p_grid,MW))+
  geom_point()+
  geom_line()+
  theme_classic()+  
  geom_line(aes(p_grid,prior/1000),color='red')+xlab("Knowledge of CogSci") + ylab("posterior probability")

attach(mtcars)
par(mfrow=c(2,2))
RF_5
KT_5
JS_5
MW_5


```


6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)?

7. Bonus knowledge: all the stuff we have done can be implemented in a lme4-like fashion using the brms package. Here is an example.
```{r}
library(brms)
d <- data.frame(
  Correct=c(3,2,160,66),
  Questions=c(6,2,198,132),
  Teacher=c("RF","KT","JS","MW"))

FlatModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("uniform(0,1)", class = "Intercept"))
plot(FlatModel)
PositiveModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.8,0.2)", class = "Intercept"))
plot(PositiveModel)
SkepticalModel <- brm(Correct|trials(Questions)~1,data=subset(d,Teacher=="RF"),prior=prior("normal(0.5,0.01)", class = "Intercept"))
plot(SkepticalModel)
```

If you dare, try to tweak the data and model to test two hypotheses:
- Is Kristian different from Josh?
- Is Josh different from chance?



